---
title: "Expected Net Points | Different Methods"
author: "Julian McClellan"
date: "June 25, 2017"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(latex2exp)
library(magrittr)

knitr::opts_chunk$set(message = FALSE, echo = FALSE)

theme_set(theme_minimal())

EXP_DAT <- read_csv("../raw_fdowns_nscore_half_and_reset.csv")
KOFF <- EXP_DAT %>%
  filter(Follow_Kickoff == 1)
```

```{r construct_vars}
# extracts yrdline stat from the tidy df
eys <- function(tidy_df){
  pull(tidy_df, -Yfog) }

# For Expected Net Points until End of Half Using V_K
V_k <- KOFF  %$%
  mean(Net_Score_to_Half)

P_o <- EXP_DAT %>% 
  mutate(offense_score = ifelse(Reset_Team_to_Score == 1, 1, 0)) %>% 
  group_by(Yfog) %>%
  summarise(P_o = mean(offense_score))

P_d <- EXP_DAT %>% 
  mutate(defense_score = ifelse(Reset_Team_to_Score == - 1, 1, 0)) %>% 
  group_by(Yfog) %>%
  summarise(P_d = mean(defense_score))

enp_o_oscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 1) %>%
  group_by(Yfog) %>%
  summarise(enp_o_oscore_yd = mean(Net_Score_to_Reset))

enp_o_dscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == -1) %>%
  group_by(Yfog) %>%
  summarise(enp_o_dscore_yd = mean(Net_Score_to_Reset))

enp_o_noscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 0) %>%
  group_by(Yfog) %>%
  summarise(enp_o_noscore_yd = mean(Net_Score_to_Reset))

# For Expected Net Points until End of Half using V_k recursive
ko_P_o <- KOFF %>%
  mutate(offense_score = ifelse(Reset_Team_to_Score == 1, 1, 0)) %$%
  mean(offense_score)

ko_P_d <- KOFF %>% 
  mutate(defense_score = ifelse(Reset_Team_to_Score == - 1, 1, 0)) %$%
  mean(defense_score)

ko_enp_o_oscore_yd <- KOFF %>%
  filter(Reset_Team_to_Score == 1) %$%
  mean(Net_Score_to_Reset)

ko_enp_o_dscore_yd <- KOFF %>%
  filter(Reset_Team_to_Score == -1) %$%
  mean(Net_Score_to_Reset)

ko_enp_o_noscore_yd <- KOFF %>%
  filter(Reset_Team_to_Score == 0) %$%
  mean(Net_Score_to_Reset)

V_k_rec <- (ko_P_o * ko_enp_o_oscore_yd +
            ko_P_d * (- ko_enp_o_dscore_yd) +
            (1 - ko_P_o - ko_P_d) * ko_enp_o_noscore_yd) /
            (1 + ko_P_o + ko_P_d)

```

# Scalar Variables Table

Variable | Value
-------- | ----------
$P(O \;\|\; KO\;1st Down)$ | `r ko_P_o`
$P(D \;\|\; KO\;1st Down)$ | `r ko_P_d`
$(ExpPTS_O \;\|\; O, KO)$ | `r ko_enp_o_oscore_yd`
$(ExpPTS_O \;\|\; D, KO)$ | `r ko_enp_o_dscore_yd`
$(ExpPTS_O \;\|\; No\;TD/FG, KO)$ | `r ko_enp_o_noscore_yd`
$V_k$ | `r V_k`
$V_{k}^rec$ | `r V_k_rec`

\pagebreak

# Expected Net Points Until End of Half

```{r}
# Joins the data in several tables with yfog stats into one
combine_yfog_stats <- function(..., col_name, melt = TRUE){
  data_frames <- list(...)
  
  for (df in data_frames){
    if (!exists("base_df")){
      base_df <- df
    } else {
      base_df <- inner_join(base_df, df)
    }
  }
  if (melt){
    base_df %>%
      select(-Yfog) %>%
      melt(value.name = col_name) %>%
      mutate(Yfog = rep(1:99, length(data_frames)))
  } else {
   base_df 
  }
}

enp_eoh_avg <- EXP_DAT %>%
  group_by(Yfog) %>%
  summarise(enp_eoh_avg = mean(Net_Score_to_Half))

enp_eoh_vk <- tibble(Yfog = 1:99,
                     enp_eoh_vk = eys(P_o) * (eys(enp_o_oscore_yd) - 
                       V_k) -  eys(P_d) * (- eys(enp_o_dscore_yd) + 
                       V_k) + (1 - eys(P_o) - eys(P_d)) * 
                       (eys(enp_o_noscore_yd))
                     )

enp_eoh_vk_rec <- tibble(Yfog = 1:99,
                     enp_eoh_vk_rec = eys(P_o) * (eys(enp_o_oscore_yd) - 
                       V_k_rec) -  eys(P_d) * (- eys(enp_o_dscore_yd) + 
                       V_k_rec) + (1 - eys(P_o) - eys(P_d)) * 
                       (eys(enp_o_noscore_yd))
                     )

enp_all <- combine_yfog_stats(enp_eoh_avg, enp_eoh_vk, enp_eoh_vk_rec, col_name = "enp")

ggplot(enp_all, aes(Yfog, enp, color = variable)) + 
  geom_line() + 
  geom_line() + 
  geom_line() + 
  scale_color_manual(name = "Net Points Type", values = c("black", "blue", "red"),
                     labels = c("Average", "V_k",
                                "V_k recursive")) + 
  labs(title = "Expected Net Points Comparison",
       x = "Yards from Own Goal",
       y = "Expected Net Points until End of Half (Offense)")
```


# Probability of Offense/Defense/No One Scoring a TD/FG Before Halftime by Yardline


```{r}
### All First and 10/Goal
one_min_o_min_d <- tibble(Yfog = 1:99, omomd = 1 - eys(P_o) - eys(P_d))
score_prob <- combine_yfog_stats(P_o, P_d, one_min_o_min_d, col_name = "score_prob")

ggplot(score_prob, aes(Yfog, score_prob, color = variable)) + 
  geom_line() + 
  geom_line() + 
  geom_line() + 
  scale_color_manual(name = "Who Scores TD/FG Before Halftime", values = c("black", "blue", "red"),
                     labels = c("Offense", "Defense",
                                "No TD/FG Before Half")) + 
  labs(title = "Prob of O/D/No One Scoring a TD/FG Before Halftime by Yardline",
       subtitle = "All First and 10s/Goals",
       x = "Yards from Own Goal",
       y = "Probability")
```

# Expected Net Points Conditional


```{r}
### All First and 10/Goal
enp_cond_all <- combine_yfog_stats(enp_o_oscore_yd, enp_o_dscore_yd, enp_o_noscore_yd, 
                                   col_name = "enp_cond")

ggplot(enp_cond_all, aes(Yfog, enp_cond, color = variable)) + 
  geom_line() + 
  geom_line() + 
  geom_line() + 
  scale_color_manual(name = "Who Scores TD/FG Before Halftime",
                     values = c("black", "blue", "red"),
                     labels = c("Offense", "Defense",
                                "No TD/FG Before Half")) + 
  labs(title = "Conditional Expected Net Points for Offense by Yard Line",
       subtitle = "All First and 10s/Goals",
       x = "Yards from Own Goal",
       y = "Expected Net Points (Offense)")
```



```{r create_csv}
# g1 col a = enp(x) normal way mean (to the half)
# g1 col b = enp(x) with (V_k ^ rec)
# g1 col c = enp(x) with V_k
# g2 col d, e = Prob(O | yrdline = x), Prob(D | yrdline = x)
# g2 col f = 1 - col d - col e
# g3 col g = (ExpNetPTS_O | O, yrdline = x)
# g3 col h = (ExpNetPTS_O | D, yrdline = x)
# g3 col i = (ExpNetPTS_O | No TD/FG, yrdline = x) 
obs_enp_eoh_avg <- EXP_DAT %>%
  group_by(Yfog) %>%
  summarise(obs_enp_eoh_avg = n())

obs_enp_o_oscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 1) %>%
  group_by(Yfog) %>%
  summarise(obs_enp_o_oscore_yd = n())

obs_enp_o_dscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == -1) %>%
  group_by(Yfog) %>%
  summarise(obs_enp_o_dscore_yd = n())

obs_enp_o_noscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 0) %>%
  group_by(Yfog) %>%
  summarise(obs_enp_o_noscore_yd = n())
output_csv <- combine_yfog_stats(enp_eoh_avg, enp_eoh_vk, enp_eoh_vk_rec, 
                                 P_o, P_d, one_min_o_min_d, enp_o_oscore_yd,
                                 enp_o_dscore_yd, enp_o_noscore_yd, 
                                 obs_enp_eoh_avg,  
                                 obs_enp_o_oscore_yd, obs_enp_o_dscore_yd,
                                 obs_enp_o_noscore_yd,
                                 col_name = NA, melt = FALSE) %>%
  rename(`ENP Normal (Average)` = enp_eoh_avg,
         `ENP V_k` = enp_eoh_vk,
         `ENP V_k Recursive` = enp_eoh_vk_rec,
         `Prob O Scores TD/FG` = P_o,
         `Prob D Scores TD/FG` = P_d,
         `1 - P(O) - P(D)` = omomd,
         `ENP_O | O, yrdline = x` = enp_o_oscore_yd,
         `ENP_O | D, yrdline = x` = enp_o_dscore_yd,
         `ENP_O | No TD/FG, yrdline = x` = enp_o_noscore_yd,
         `# Obs ENP Normal (Average)` = obs_enp_eoh_avg,
         `# Obs (ENP_O | O, yrdline = x)` = obs_enp_o_oscore_yd,
         `# Obs (ENP_O | D, yrdline = x)` = obs_enp_o_dscore_yd,
         `# Obs (ENP_O | No TD/FG, yrdline = x)` = obs_enp_o_noscore_yd
         ) %>%
  write_csv("enp_methods_output.csv")
```

# Appendix: Variable Construction Formulas

$$
V_{k} ^ {rec} = \frac{Prob(O\; |\;KO) * ExpPTS_O + Prob(D\;|\;KO) * (- ExpPTS_O) + 
[1 - Prob(O\;|\;KO) - Prob(D\;|\;KO)] * ExpNetPTS}{1}
$$

$$
ENP(yrdline = x) = Prob(O\;|\; yrdline = x) * [(ExpNetPTS_O\; | O\;Score, yrdline = x) - V_k] -
$$
$$
 Prob(D | yrdline = x) * [(-ExpPTS_O\; | D\;Score, yrdline = x) - V_k] + 
$$
$$
[1 - Prob(O|yrdline = x) - Prob(D|yrdline=x)] * (ExpNetPts_O | Not\;TD,\;Not\;FG,yrdline = x)
$$

Similarly except replace $V_k$ with $V_{k} ^ rec$.
