---
title: "Expected Points | Different Methods"
author: "Julian McClellan"
date: "June 12, 2017"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(latex2exp)

knitr::opts_chunk$set(message = FALSE)

theme_set(theme_minimal())

EXP_DAT <- read_csv("../raw_fdowns_nscore_half_and_reset.csv")
KOFF <- EXP_DAT %>%
  filter(Follow_Kickoff == 1)
```

```{r construct_vars}
# extracts yrdline stat from the tidy df
eys <- function(tidy_df){
  pull(tidy_df, -Yfog) }

# For Expected Net Points until End of Half Using V_K
V_k <- KOFF  %>%
  group_by(Yfog)  %>%
  summarise(V_k = mean(Net_Score_to_Half))

P_o <- EXP_DAT %>% 
  mutate(offense_score = ifelse(Reset_Team_to_Score == 1, 1, 0)) %>% 
  group_by(Yfog) %>%
  summarise(P_o = mean(offense_score))

P_d <- EXP_DAT %>% 
  mutate(defense_score = ifelse(Reset_Team_to_Score == - 1, 1, 0)) %>% 
  group_by(Yfog) %>%
  summarise(P_d = mean(defense_score))

enp_o_oscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 1) %>%
  group_by(Yfog) %>%
  summarise(enp_o_oscore_yd = mean(Net_Score_to_Reset))

enp_o_dscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == -1) %>%
  group_by(Yfog) %>%
  summarise(enp_o_dscore_yd = mean(Net_Score_to_Reset))

enp_o_noscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 0) %>%
  group_by(Yfog) %>%
  summarise(enp_o_noscore_yd = mean(Net_Score_to_Reset))

# For Expected Net Points until End of Half using V_k recursive
ko_P_o <- KOFF %>%
  mutate(offense_score = ifelse(Reset_Team_to_Score == 1, 1, 0)) %>% 
  group_by(Yfog) %>%
  summarise(ko_P_o = mean(offense_score))

ko_P_d <- KOFF %>% 
  mutate(defense_score = ifelse(Reset_Team_to_Score == - 1, 1, 0)) %>% 
  group_by(Yfog) %>%
  summarise(ko_P_d = mean(defense_score))

ko_enp_o_oscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 1) %>%
  group_by(Yfog) %>%
  summarise(ko_enp_o_oscore_yd = mean(Net_Score_to_Reset))

ko_enp_o_dscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == -1) %>%
  group_by(Yfog) %>%
  summarise(ko_enp_o_dscore_yd = mean(Net_Score_to_Reset))

ko_enp_o_noscore_yd <- EXP_DAT %>%
  filter(Reset_Team_to_Score == 0) %>%
  group_by(Yfog) %>%
  summarise(ko_enp_o_noscore_yd = mean(Net_Score_to_Reset))

V_k_rec <- tibble(Yfog = 1:99,
                  V_k_rec = (eys(ko_P_o) * eys(ko_enp_o_oscore_yd) -
                    eys(ko_P_d) * eys(ko_enp_o_dscore_yd) +
                    (1 - eys(ko_P_o) - eys(ko_P_d)) * eys(ko_enp_o_noscore_yd)) /
                    (1 + eys(ko_P_o) + eys(ko_P_d))
                  ) 
```

# Expected Net Points Until End of Half

```{r}
# Joins the data in several tables with yfog stats into one
combine_yfog_stats <- function(..., col_name){
  data_frames <- list(...)
  
  for (df in data_frames){
    if (!exists("base_df")){
      base_df <- df
    } else {
      base_df <- inner_join(base_df, df)
    }
  }
  base_df %>%
    select(-Yfog) %>%
    melt(value.name = col_name) %>%
    mutate(Yfog = rep(1:99, length(data_frames)))
}

 foreoh_avg <- EXP_DAT %>%
  group_by(Yfog) %>%
  summarise(enp_eoh_avg = mean(Net_Score_to_Half))

enp_eoh_vk <- tibble(Yfog = 1:99,
                     enp_eoh_vk = eys(P_o) * (eys(enp_o_oscore_yd) - 
                       eys(V_k)) -  eys(P_d) * (- eys(enp_o_dscore_yd) + 
                       eys(V_k)) + (1 - eys(P_o) - eys(P_d)) * 
                       (eys(enp_o_noscore_yd))
                     )

enp_eoh_vk_rec <- tibble(Yfog = 1:99,
                         enp_eoh_vk_rec = eys(ko_P_o) * (eys(ko_enp_o_oscore_yd) 
                           -  eys(V_k_rec)) -  eys(ko_P_d) * 
                          (- eys(ko_enp_o_dscore_yd) +  eys(V_k_rec)) + 
                          (1 - eys(ko_P_o)  - eys(ko_P_d)) *
                          (eys(ko_enp_o_noscore_yd))
                        )

enp_all <- combine_yfog_stats(enp_eoh_avg, enp_eoh_vk, enp_eoh_vk_rec, col_name = "enp")

ggplot(enp_all, aes(Yfog, enp, color = variable)) + 
  geom_line() + 
  geom_line() + 
  geom_line() + 
  scale_color_manual(name = "Net Points Type", values = c("black", "blue", "red"),
                     labels = c("Average", "V_k",
                                "V_k recursive")) + 
  labs(title = "Expected Net Points Comparison",
       x = "Yards from Own Goal",
       y = "Expected Net Points until End of Half (Offense)")
```


# Probability of Offense/Defense/No One Scoring a TD/FG Before Halftime by Yardline

### Any First and 10/Goal

```{r}
one_min_o_min_d <- tibble(Yfog = 1:99, omomd = 1 - eys(P_o) - eys(P_d))
score_prob <- combine_yfog_stats(P_o, P_d, one_min_o_min_d, col_name = "score_prob")

ggplot(score_prob, aes(Yfog, score_prob, color = variable)) + 
  geom_line() + 
  geom_line() + 
  geom_line() + 
  scale_color_manual(name = "Who Scores TD/FG Before Halftime", values = c("black", "blue", "red"),
                     labels = c("Offense", "Defense",
                                "No TD/FG Before Half")) + 
  labs(title = "Prob of Offense/Defense/No One Scoring a TD/FG Before Halftime by Yardline",
       subtitle = "All First and 10s/Goals",
       x = "Yards from Own Goal",
       y = "Probability")
```

### Only First and 10/Goal Following Kickoff

```{r}
ko_one_min_o_min_d <- tibble(Yfog = 1:99, omomd = 1 - eys(ko_P_o) - eys(ko_P_d))
ko_score_prob <- combine_yfog_stats(ko_P_o, ko_P_d, ko_one_min_o_min_d, col_name = "ko_score_prob")

ggplot(ko_score_prob, aes(Yfog, ko_score_prob, color = variable)) + 
  geom_line() + 
  geom_line() + 
  geom_line() + 
  scale_color_manual(name = "Who Scores TD/FG Before Halftime", values = c("black", "blue", "red"),
                     labels = c("Offense", "Defense",
                                "No TD/FG Before Half")) + 
  labs(title = "Prob of Offense/Defense/No One Scoring a TD/FG Before Halftime by Yardline",
       subtitle = "First and 10s/Goals Following a Kickoff",
       x = "Yards from Own Goal",
       y = "Probability")
```

g1 col a = enp(x) normal way mean (to the half)
g1 col b = enp(x) with (V_k ^ rec)
g1 col c = enp(x) with V_k
g2 col d, e = Prob(O | yrdline = x), Prob(D | yrdline = x)
g2 col f = 1 - col d - col e
g3 col g = (ExpNetPTS_O | O, yrdline = x)
g3 col h = (ExpNetPTS_O | D, yrdline = x)
g3 col i = (ExpNetPTS_O | No TD/FG, yrdline = x) 
g4 col j = V_k ^ rec
g4 col k = V_k

# Expected Net Points for the Offense | Offense Scored TD/FG, Defense Did, No One Did

## Variable Construction

$$V_k = Expected\; Net\; Points \; from \; Kickoff \; until\; EOH$$

$$
V_{k} ^ {rec} = \frac{Prob(O\; |\;KO) * ExpPTS_O + Prob(D\;|\;KO) * ExpPTS_D + 
[1 - Prob(O\;|\;KO) - Prob(D\;|\;KO)] * ExpNetPTS}{1}
$$

$$Prob(O\; |\; yrdline = x)$$


$$Prob(D\; |\; yrdline = x)$$


$$ExpNetPTS_o\; |\; O\; Score,\; yrdline = x$$

```{r}
```

$$ExpNetPTS_d\; |\; D\; Score,\; yrdline = x$$

$$ExpNetPTS_O\; |\; D\; Score,\; yrdline = x$$

```{r}

```

$$ExpNetPTS_O\; |\; No\; TD,\; No\; FG,\; yrdline = x$$

```{r}
```

## Expected Net Points ($V_k$)

$$
ENP(yrdline = x) = Prob(O\;|\; yrdline = x) * [(ExpNetPTS_O\; | O\;Score, yrdline = x) - V_k] -
$$
$$
 Prob(D | yrdline = x) * [(-ExpPTS_O\; | D Score, yrdline = x) - V_k] + 
$$
$$
[1 - Prob(O|yrdline = x) - Prob(D|yrdline=x)] * (ExpNetPts_O | Not\;TD,\;Not\;FG,yrdline = x)
$$

```{r}
```

