---
title: "Method 2 | Regression Approach"
author: "Julian McClellan"
date: "August 11th Data"
output: 
  pdf_document:
    toc: true
---
# Expected Net Points

```{r setup, include=T, message=F, warning=F, echo=F, fig.width=7.5, fig.height=10}
knitr::opts_chunk$set(echo = F, message = F, warning = F, include = T)
library(tidyverse)
library(ggplot2)
library(reshape2)
library(glue)

theme_set(theme_minimal())

all <- read.csv("M2_regression_all.csv")
nn <- read.csv("M2_regression_99.csv")
var <- read.csv("M2_regression_var.csv")

cols_of_int = c(
        "v99_iv_clust_2000.2016",
         "v99_iv_clust_2000.2010",
         "v99_iv_clust_2011.2015",
         "v99_iv_clust_2016.2016"
         )

v_iv <- nn %>%
  select(v99_iv_clust_2000.2016,
         v99_iv_clust_2000.2010,
         v99_iv_clust_2011.2015,
         v99_iv_clust_2016.2016
         ) %>%
  mutate(Yfog = 1:99) %>%
  melt(id.var = "Yfog", variable.name = "split", measure.vars = cols_of_int,
       value.name = "v99_iv")

v_iv <- v_iv %>%
  mutate(full_or_not = ifelse(grepl("2000.2016", split), "", " "))

(
  ggplot(v_iv) +
  geom_line(aes(x = Yfog, y = v99_iv, color = split)) +
  labs(x = "Yards from Own Goal",
       y = "Expected Net Points Until Half",
       title = "Expected Net Points Until Half",
       subtitle = "Regression Approach IV | Method 2",
       color = "Season Split") +
  scale_color_manual(labels = c("2000 - 2016", "2000 - 2010",
                                "2011 - 2015", "2016"),
                     values = c("red", "black", "blue", "green")) +
  facet_grid(full_or_not ~ .)
  )
```


```{r, fig.width=7.5, fig.height=10}
v_ols <- all %>%
  select(v_ols_clust_2000.2016,
         v_iv_clust_2000.2016)

v_ols <- v_ols[-100,] %>% 
  mutate(Yfog = 1:99) %>%
  melt(id.var = "Yfog", variable.name = "split", measure.vars = 
         c("v_ols_clust_2000.2016", "v_iv_clust_2000.2016"), value.name = "v99_iv_ols") %>%
  mutate(iv_or_not = ifelse(grepl("iv", split), "", " "))

(
  ggplot(v_ols) +
  geom_line(aes(x = Yfog, y = v99_iv_ols, color = split)) +
  labs(x = "Yards from Own Goal",
       y = "Expected Net Points Until Half",
       title = "Expected Net Points Until Half",
       subtitle = "Regression Approach IV versus OLS | Method 2",
       color = "IV or OLS") +
  scale_color_manual(labels = c("2000 - 2016 OLS",
                                "2000 - 2016 IV"),
                     values = c("black", "red")) +
  facet_grid(iv_or_not ~ .)
  )
```

# Standard Errors
```{r fig.width=7.5, figh.height=5}
cols_of_int = c(
        "se_v_iv_clust_2000.2016",
         "se_v_iv_clust_2000.2010",
         "se_v_iv_clust_2011.2015",
         "se_v_iv_clust_2016.2016"
         )
se_g1 <- all %>%
  select(se_v_iv_clust_2000.2016,
         se_v_iv_clust_2000.2010,
         se_v_iv_clust_2011.2015,
         se_v_iv_clust_2016.2016)

se_g1 <- se_g1[-100, ] %>%
  mutate(Yfog = 1:99) %>%
  melt(id.var = "Yfog", variable.name = "split", measure.vars = cols_of_int,
       value.name = "se_g1") %>%
  mutate(full_or_not = ifelse(grepl("2000.2016", split), "", " "))

(
  ggplot(se_g1) +
  geom_point(aes(x = Yfog, y = se_g1, color = split), size = .8) +
  labs(x = "Yards from Own Goal",
       y = "Standard Error",
       title = "Standard Error | Method 2 Regression",
       subtitle = "Subsample Comparison",
       color = "Legend") +
  scale_color_manual(labels = c("2000 - 2016 IV Half", "2000 - 2010 IV Half",
                                "2011 - 2015 IV Half", "2016 IV Half"),
                     values = c("red", "black", "blue", "green")) +
  facet_grid(full_or_not ~ .)
  )



g3_inp_list <- list(iv = c(T, T, T),
       clust = c(T, T, T),
       cluster_var = c("Armchair_gid", "Half_num", "Armchair_pid"),
       source_dir = "expected_points/method_2/",
       year_start = c(2000, 2000, 2000),
       year_stop = c(2016, 2016, 2016),
       all_or_nn = "nn"
   ) 


# Only calling g3 and g4 because google spreadsheet notes list
# them as separate graphs, however I think they should be together (g3 top, g4
# bottom)

for_table <- read.csv("M2_g3_se_99.csv")

g3_cols <-c("se_v_iv_clust_Armchair_gid_2000.2016",
         "se_v_iv_clust_Half_num_2000.2016",
         "se_v_iv_no_clust_Armchair_pid_2000.2016") 

g3_se <- read.csv("M2_g3_se_99.csv") %>%
  select(se_v_iv_clust_Armchair_gid_2000.2016,
         se_v_iv_clust_Half_num_2000.2016,
         se_v_iv_no_clust_Armchair_pid_2000.2016) %>%
  mutate(g3 = "") 

all <- all %>%
  bind_cols(as_tibble(g3_se$se_v_iv_clust_Armchair_gid_2000.2016))

write_csv(all, "M2_regression.csv")

g3_se <- g3_se[-100,] %>%
  mutate(Yfog = 1:99) %>%
  melt(id.var = "Yfog", variable.name = "split", measure.vars = g3_cols,
       value.name = "g34_se") %>%
  mutate(g3 = "") # For use in faceting

g4_cols <- c("se_v_iv_clust_2000.2016", "se_v_ols_no_clust_2000.2016")

g4_se <- all %>%
  select(se_v_iv_clust_2000.2016,
         se_v_ols_no_clust_2000.2016) 

g4_se <- g4_se[-100,] %>% mutate(Yfog = 1:99) %>%
  melt(id.var = "Yfog", variable.name = "split", measure.vars = g4_cols,
       value.name = "g34_se") %>%
  mutate(g3 = " ")

g34_se <- rbind(g3_se, g4_se)

(
  ggplot(g34_se) +
  geom_point(aes(x = Yfog, y = g34_se, color = split), size = .8, alpha = .5) +
  labs(x = "Yards from Own Goal",
       y = "Standard Error",
       title = "Standard Error | Method 2 Regression",
       subtitle = "IV versus OLS",
       color = "Legend") +
  scale_color_manual(labels = c("2000 - 2016 IV Game",
                                "2000 - 2016 IV Half",
                                "2000 - 2016 IV None",
                                "2000 - 2016 IV Half",
                                "2000 - 2016 OLS None"),
                     values = c("purple", "red", "orange", "red", "black")) +
  facet_grid(g3 ~ .)
  )
```

\newpage

\begin{table}[]
\centering
\caption{v100}
\label{my-label}
\begin{tabular}{lllll}
Technique & Cluster & years     & v100\_coef & v100\_se \\
IV         &  half     &  2000-2016  & `r all$v_iv_clust_2000.2016[100]` &  `r all$se_v_iv_clust_2000.2016[100]`   \\
IV         &  half     &  2000-2010  & `r all$v_iv_clust_2000.2010[100]` &  `r all$se_v_iv_clust_2000.2010[100]` \\
IV         &  half     &  2011-2015  & `r all$v_iv_clust_2011.2015[100]` &  `r all$se_v_iv_clust_2011.2015[100]`         \\
IV         &  half     &  2016       & `r all$v_iv_clust_2016.2016[100]` &  `r all$se_v_iv_clust_2016.2016[100]`          \\
IV         &  none     &  2000-2016  & `r all$v_iv_no_clust_2000.2016[100]` &  `r all$se_v_iv_no_clust_2000.2016[100]`          \\
IV         &  game     &  2000-2016  & `r for_table$v_iv_clust_Armchair_gid_2000.2016[100]` &  `r for_table$se_v_iv_clust_Armchair_gid_2000.2016[100]`          \\
OLS        &  none     &  2000-2016  & `r all$v_ols_no_clust_2000.2016[100]` &  `r all$se_v_ols_no_clust_2000.2016[100]` 
\end{tabular}
\end{table}